Figures for Paper

### 2021-06-30 paper_figures.Rmd
add wtc11 snp cleaner

<!-- # iPSC WTC11 parent line genotype -->

<!-- ```{r confirm_no_als_ftd_wtc11 mutations} -->
<!-- library(vcf) -->
<!-- wtc11_vcf = "/Users/annaleigh/Downloads/3546dc62_AH77TTBBXX_DS-229105_GCCAAT_recalibrated.vcf.gz" -->
<!-- vcf <- read.vcfR( wtc11_vcf, verbose = FALSE ) -->
<!-- ``` -->

@@

clean_data_table %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = unc13a_detected,y = log10(library_size) )) + 
  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
  labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue + disease_group2) + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
  labs(y = "log10 library size", x = "UNC13A CE detected") +
  theme(axis.text.x = element_text(size = 15)) + 
  ggpubr::theme_pubr()
```

@@

clean_data_table %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = unc13a_detected,y = UNC13A_TPM )) + 
  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
  labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue + disease_group2, scales = 'free_y') + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
  labs(y = "UNC13A expression (TPM)", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr()

@@

```{r unc13a tpm and ce detection}
## effect of RIN?
clean_data_table %>%
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  add_count(tissue_clean,disease_group2) %>% 
  mutate(plot_name = glue::glue("{disease_group2} \n ({n})")) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = plot_name, fill = unc13a_detected,y = UNC13B_TPM )) + 
  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
  geom_boxplot(outlier.colour = NA) +
  labs(title = "UNC13A CE detection and UNC13B TPM",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue_clean,scales = 'free') + 
  ggpubr::stat_compare_means(label = "p.signif",) +
  labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr()
```

@@

cell_types %>% 
  left_join(clean_data_table %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ))

>>>>

@@

clean_data_table %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
  mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
  ggplot(aes(x = disease_group2,y = log10(library_size), fill = unc13a_detected)) + 
  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
  geom_boxplot(colour = "black", outlier.colour = NA) +
  labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue, scales = 'free_x') + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
  labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
  theme(axis.text.x = element_text(size = 15)) + 
  ggpubr::theme_pubr()
```

@@

clean_data_table %>%
    filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
  mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
  ggplot(aes(x = disease_group2,y = UNC13A_TPM, fill = unc13a_detected)) + 
  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
  geom_boxplot(colour = "black", outlier.colour = NA) +
  labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue, scales = 'free') + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
  labs(y = "UNC13A TPM", fill= "UNC13A CE detected", x = element_blank()) +
  theme(axis.text.x = element_text(size = 15)) + 
  ggpubr::theme_pubr()

@@

## Effect of UNC13A discovery and UNC13B TPM

```{r unc13a ce and UNC13B_TPM}
## effect of RIN?
clean_data_table %>%
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  add_count(tissue_clean,disease_group2) %>% 
  mutate(plot_name = glue::glue("{disease_group2} \n ({n})")) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = plot_name, fill = unc13a_detected,y = UNC13B_TPM )) + 
  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
  geom_boxplot(outlier.colour = NA) +
  labs(title = "UNC13A CE detection and UNC13B TPM",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue_clean,scales = 'free') + 
  ggpubr::stat_compare_means(label = "p.signif",) +
  labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr()
```

@@

cell_types %>% 
  left_join(clean_data_table, by = 'sample') %>%
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
  filter(!is.na(inclusion_reads)) %>% 
  select(sample,cell,deconv, tissue_clean,unc13a_detected) %>% 
  unique() %>% 
  ggplot() + 
  geom_boxplot(aes(x = unc13a_detected, y = deconv, fill = cell)) + 
  facet_wrap(~ tissue_clean,scales = 'free')
